---
title: "ST661: Assignment 1"
author: "Alok Kumar Singh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Question 1
a.
```{r}
x<-10:14
x<-c(10,11,12,13,14)
x<-seq(10,14, by = 1)
x
```
b.
```{r}
y<- c(14,10,11,13,12)
y
```
c.
```{r}
x[3]
```

d.
```{r}
y[c(2,4)]
```

e.
```{r}
sum(x>y)
```
Sum(x>y) will give total number of values in x and y vector where value in x vector at a particular index is greater then the value of y vector at that index.

f.
```{r}
sum(abs(x-y)==1)
```
sum(abs(x-y)==1) will give the total number of times after comparing the absolute value of x at particular index and subtracting it with y at that index to 1. 
g.
```{r}
a<-(x^2)
b<-sum(a)/sum(y)
b
```
h.
```{r}
cbind(x,y)
```


#### Question 2
a.
```{r}
house <- read.csv("house.csv")
head(house)
```
b.
```{r}
subset(house,Country == 'Ireland')
```
c.
```{r}
f<-subset(house,Country == 'Spain')
f$TooDark
```

d.
```{r}
g<-subset(house,NoIndoorToilet == 0)
g
```

e.
```{r}
h<-subset(house,NoIndoorToilet == 0 & NoBathShower == 0)
h
```


f.
```{r}
h$Country
```

g.
```{r}
max(house$Damp)
```



#### Question 3
a.
```{r}
load("worldcup.Rdata")
barplot(ngoals)
mean1 <- mean(ngoals)
print(paste("Mean is-",mean1))
median1 <- median(ngoals)
print(paste("Median is-",median1))
uniqv<-unique(ngoals)
mode1<-uniqv[which.max(tabulate(match(ngoals, uniqv)))]
print(paste("The most frequent number of goals is -",mode1))
```
b.
```{r}
vg <-ngoals[c(52,104,168,232)]
vg
```
c.
```{r}
year<-rep(seq(1990,2002, by = 4),times = c(52,52,64,64))
year
```

d.
```{r}
factor(year)
df <- data.frame(ngoals,year)
df
```

e.
```{r}
matchn<-rep(c(1:52,1:52,1:64,1:64))
matchn
df1 <- data.frame(df,matchn)
df1
```


f.
```{r}
df2 <- df1[c(52,104,168,232),]
df2
df2[which.max(df2$ngoals),2]

```

g.
```{r}
boxplot(ngoals~year,col = c("red","green","yellow","orange"))
```

The goal 8 is outlier as it falls outside the IQR which is 6 for this plot.

h.
```{r}

```
